---
title: "p8105 Homework 1"
author: "Tory Lynch"
date: "9/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tibble)
library(dplyr)
```

##Problem 1

**Part 1**

I cannot calculate the mean of each variable because the character and factor variables are not numbers, and therefore cannot be averaged as number. The mean of x is 2.446 and the mean of the logical variable is 0.7, which indicates taht x is greater than 2 in 70% of the runs. 

```{r}
set.seed(1234)

#Create dataframe 
problem_one_df <- tibble(
  x = runif(n = 10, min = 0, max = 5),
  prob_one_logical = ifelse(x > 2, TRUE, FALSE),
  prob_one_character = c("data", "science", "home", "work", "number", "one", "friday", "september", "21", "2018"),
  prob_one_factor = factor(c("fall", "winter", "spring", "summer", "fall", "fall", "winter", "winter", "fall", "summer"))
) 

#Show dataframe table
problem_one_df

#Find mean for each column. 
  #colMeans(problem_one_df) - commented out so that the chunk can run smoothly. 
  colMeans(problem_one_df[1:2])

```

**Part 2**

I am now able to calculate the mean of the factor variable after converting it into a numeric variable; each factor was assigned a numeric value (1 = fall, 2 = spring etc.) and the average is 2.4. This value is not helpful for interpretation, since the numbers were not assigned in "order" of the seasons (i.e. fall is 1 and winter is 4). The non-number character variables were assigned as "NAs" so while I cannot calculate a mean value for them, they do not inhibit the code and are assigned "NA" in the table of means. The means of x and the logical variable are unchanged. 

```{r}

#Convert character, factor, and logical variables to numeric variables 
problem_one_df <- problem_one_df %>% mutate_all(funs(as.numeric))

#Find mean for each column as numerical variables 
means_table <-colMeans(problem_one_df)

#Show dataframe table with converted variables
problem_one_df

```

**Part 3**

The converted factor variable (factor --> character --> numeric) is the same as in part 2 in that each level of the factor (i.e. each season) is assigned the same numeric value. The converted character variable differs from that in part 2 because the two character values that were numbers are converted into factors; so, instead of the values being "21" and "2018", they are converted to 1 and 2, respectivley, becasue they are categorized as two different factors. The rest of the character values (the words) remain as NAs. 

```{r}

new_numeric_one <-  as.numeric(as.factor(problem_one_df$prob_one_character))
new_numeric_one

new_numeric_two <- as.numeric(as.character(problem_one_df$prob_one_factor))
new_numeric_two
```